### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@34b8a1224cd30f556e77c6ce7ab7ff78d85122b4 🚀

\begin{figure}
\hypertarget{fig:biogeo}{%
\centering
\includegraphics{figures/combined_biogeo.png}
\caption{Phylogeographic regions of bats (top) and viruses (bottom)
based on the joint analysis of their occurrence and phylogenetic
relatedness. The different colors show tendencies to separate alongside
the first two components of a PCoA. Note that the PCoA for the bats and
viruses are independent, and so cannot be compared directly -- that
being said, the regions can be compared across maps.}\label{fig:biogeo}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:richness}{%
\centering
\includegraphics{figures/combined_richness.png}
\caption{Top panel: relative diversity of known bat hosts of
betacoronaviruses. This map shows that the region with the largest
number of possible hosts is South-Eastern Asia. Bottom panel: congruence
between the evolutionary distinctiveness of the hosts (grey to blue) and
the viruses (grey to red).}\label{fig:richness}
}
\end{figure}

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
differential coevolutionary processes; in turn, the coevolutionary
mosaic contributes to heterogeneity in emergence risk. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, these
ideas point to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts---and their diverse immune
strategies---\textsuperscript{23}creates a landscape of selective
pressure; the trajectory of viruses' coevolutionary response is, in
turn, constrained by the opportunities they have for either
coevolutionary specialization or diversification through host jumps and
recombination.

Based on the geographic mosaic theory of coevolution, we developed a
trivariate map of three facets of coevolutionary pressure (see Methods):
(1) \emph{host phylogenetic diversity}: a high diversity of evolutionary
histories should expose viruses to more variation in host immune traits;
(2) \emph{host community uniqueness}: exposure to greater host trait
heterogeneity can drive viral diversification, and coevolving with more
unique host communities should create more unique branches of viral
evolution; and (3) propensity for \emph{viral sharing}: frequent
cross-species transmission may act as a buffer on selective pressure,
while lower rates of exchange may enable more simultaneous trajectories
of viral specialization to coexist within a given community. We combine
global maps of all three to generate a map of coevolutionary regimes,
where close colors represent similar risks, and paler pixels represent
overall higher risk (fig.~\ref{fig:trivariate}). We find that these
regions do not neatly overlap with those defined in
fig.~\ref{fig:biogeo} or fig.~\ref{fig:richness}, reinforcing the notion
that local-scale coevolutionary mosaics can form within cophylogenetic
regions.

\begin{figure}
\hypertarget{fig:trivariate}{%
\centering
\includegraphics{figures/risk_trivariate.png}
\caption{Trivariate additive mapping of the components of risk. Viral
sharing runs from yellow (low) to blue (high); host phylogenetic
diversity runs from pink (low) to high (green) channel, and
compositional uniqueness runs from cyan (low) to red (high). The GMTC
suggests that the highest evolutionary potential for emergence exists in
unique and diverse host communities with low viral sharing, \emph{i.e.}
pixels around yellow. All components are scaled in brightness so that a
pixel with no sharing, no phylogenetic diversity, and no compositional
uniqueness would be black, and a pixel with maximal values for each
would be white.}\label{fig:trivariate}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:risk}{%
\centering
\includegraphics{figures/risk_map.png}
\caption{Evolutionary potential for zoonotic emergence of bat-origin
betacoronaviruses. Risk is a composite measure of the color value and
angular distance to the yellow hue in fig.~\ref{fig:trivariate} (see
Methods).}\label{fig:risk}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:compound}{%
\centering
\includegraphics{figures/risk_compounded.png}
\caption{Overlap of the percent of each pixel occupied by urbanized
structures, representing the degree of settlement, on the spillover risk
map (where the risk comes only from wildlife, and ignores multi-hosts
chains of transmissions including non-bats hosts). Darker pixels
correspond to more risk, in that the GMTC-derived risk of
fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
human populations. This approach increases the relative risk of several
regions in Africa, and highlights the risk in India, southeast China,
and the Arabian peninsula where areas of high to moderate risk overlap
with areas of denser population.}\label{fig:compound}
}
\end{figure}

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{38,39} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species.\textsuperscript{40,41}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to human.\textsuperscript{42} For example,
bats are an essential component of many seed-dispersal
networks.\textsuperscript{43} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{44,45} and
vectors of pathogens that put a risk on human
health.\textsuperscript{46,47} Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses,\textsuperscript{18} bat
coronaviruses,\textsuperscript{2} and even specifically
betacoronaviruses\textsuperscript{19} in both the Amazon and southeast
Asia. While we find that both regions are characterized by highly
divergent host and viral communities, our framework identifies key
differences between the regions. We find that Latin America is a hotspot
of both host and viral distinctiveness, suggesting that this branch of
the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin
America.\textsuperscript{2} In contrast, in southeast Asia, host
richness and viral distinctiveness are high but sharing is low; this
suggests a different type of evolutionary dynamics that could generate
high local diversity of viruses through host switching and viral
recombination (see \emph{e.g.},\textsuperscript{13} as well as the
discovery of recombinant viruses that share genetic material from both
the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{48} Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{19,20} However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region.\textsuperscript{49}

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{50}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{51} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in humans-hosts contacts also increase the risk of emergence of
novel diseases,\textsuperscript{52} so does the changes in landscape
connectivity at local/regional scales.\textsuperscript{53} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{54} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

@@

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
\emph{PLOS ONE} \textbf{6}, e17395 (2011).}

\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
\CSLLeftMargin{44. }
\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
a Changing World}. (Springer International Publishing, 2016).
doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}

\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
\textbf{320}, 70--70 (2008).}

\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
\emph{PLOS ONE} \textbf{8}, e77183 (2013).}

\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
Research} \textbf{40}, 10--24 (2013).}

>>>>

\begin{figure}
\hypertarget{fig:biogeo}{%
\centering
\includegraphics{figures/combined_biogeo.png}
\caption{\textbf{Bat and betacoronavirus biogeographic regions.}
Phylogeography of bats (top) and viruses (bottom) is categorized based
on analysis of bat distributions paired with bat or virus phylogeny. The
different colors show tendencies to separate alongside the first two
components of a PCoA. Note that the PCoA for the bats and viruses are
independent, and so cannot be compared directly -- that being said, the
regions can be compared across maps.}\label{fig:biogeo}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:richness}{%
\centering
\includegraphics{figures/combined_richness.png}
\caption{\textbf{Bat and betacoronavirus diversity.} Top panel: relative
diversity of known bat hosts of betacoronaviruses. This map shows that
the region with the largest number of possible hosts is South-Eastern
Asia. Bottom panel: congruence between the evolutionary distinctiveness
of the hosts (grey to blue) and the viruses (grey to
red).}\label{fig:richness}
}
\end{figure}

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, the coevolutionary
mosaic underlies the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, these
ideas point to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{23} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either coevolutionary specialization or diversification through host
jumps and recombination.

Based on the geographic mosaic theory of coevolution, we developed a
trivariate map of three facets of coevolutionary pressure
(fig.~\ref{fig:trivariate}): (1) \emph{host phylogenetic diversity}: a
high diversity of evolutionary histories should expose viruses to more
variation in host immune traits; (2) \emph{host community uniqueness}:
exposure to greater host trait heterogeneity can drive viral
diversification, and coevolving with more unique host communities should
create more unique branches of viral evolution; and (3) propensity for
\emph{viral sharing}: frequent cross-species transmission may act as a
buffer on selective pressure, while lower rates of exchange may enable
more simultaneous trajectories of viral specialization to coexist within
a given community. We combine global maps of all three to generate a map
of coevolutionary regimes, where close colors represent similar risks,
and paler pixels represent overall higher risk (see Methods). We find
that these regions do not neatly overlap with those defined in
fig.~\ref{fig:biogeo} or fig.~\ref{fig:richness}, reinforcing the notion
that local-scale coevolutionary mosaics can form within cophylogenetic
regions.

\begin{figure}
\hypertarget{fig:trivariate}{%
\centering
\includegraphics{figures/risk_trivariate.png}
\caption{\textbf{Trivariate additive mapping of the components of risk.}
Viral sharing runs from yellow (low) to blue (high); host phylogenetic
diversity runs from pink (low) to high (green); and host compositional
uniqueness runs from cyan (low) to red (high). The GMTC suggests that
the highest evolutionary potential for emergence exists in unique and
diverse host communities with low viral sharing, \emph{i.e.} pixels
around yellow. All components are scaled in brightness so that a pixel
with no sharing, no phylogenetic diversity, and no compositional
uniqueness would be black, and a pixel with maximal values for each
would be white.}\label{fig:trivariate}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:risk}{%
\centering
\includegraphics{figures/risk_map.png}
\caption{\textbf{Evolutionary potential for zoonotic emergence of
bat-origin betacoronaviruses.} Risk is a composite measure of the color
value and angular distance to the yellow hue in
fig.~\ref{fig:trivariate} (see Methods).}\label{fig:risk}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:compound}{%
\centering
\includegraphics{figures/risk_compounded.png}
\caption{\textbf{Overlap between evolutionary potential and ecological
opportunity for zoonotic emergence.} Overlap of the percent of each
pixel occupied by urbanized structures, representing the degree of
settlement, on the spillover risk map (where the risk comes only from
wildlife, and ignores multi-hosts chains of transmissions including
non-bats hosts). Darker pixels correspond to more risk, in that the
GMTC-derived risk of fig.~\ref{fig:risk} is high \emph{and} the pixel is
densely occupied by human populations.}\label{fig:compound}
}
\end{figure}

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{38,39} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with more
than 1,400 estimated species.\textsuperscript{40,41} They exhibit a
broad variety of habitat use, behaviour, and feeding strategies, putting
them at key positions in the delivery and provisioning of several
ecosystem services, tied to important ecosystem-derived benefits to
humans.\textsuperscript{42} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{43,44} and
vectors of pathogens that put a risk on human
health;\textsuperscript{45,46} other species are essential links in many
seed-dispersal networks.\textsuperscript{47} Areas where bats, viruses,
and humans co-occur are not necessarily hotspots of risk for human
heath; as such, developing more precise ways to map zoonotic hazards can
help bats and humans coexist safely and support the conservation of
these important and unique animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{19} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{48}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{49}

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{50}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{51} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in human-wildlife contact also increase the risk of emergence
of novel diseases,\textsuperscript{52} as do changes in landscape
connectivity at local/regional scales.\textsuperscript{53} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{54} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

@@

\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
\CSLLeftMargin{43. }
\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
a Changing World}. (Springer International Publishing, 2016).
doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}

\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
\textbf{320}, 70--70 (2008).}

\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
\emph{PLOS ONE} \textbf{8}, e77183 (2013).}

\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
Research} \textbf{40}, 10--24 (2013).}

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
\emph{PLOS ONE} \textbf{6}, e17395 (2011).}
